Experience Report: Distributed Computing for the Common Man

### 2010-03-24 doc/icfp-2010/main.tex
*** empty log message ***

\newcommand{\Ocaml}{OCaml}
\newcommand{\unix}{\textsc{Unix}}

@@

\begin{abstract}
This is the text of the abstract.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

\terms
term1, term2

\keywords
keyword1, keyword2

@@

\section{Future Work}\label{sec:future}

A couple of caveats are in order. (1) We do not focus on issues like
data-locality or a distributed file system like Google's, rather we
assume a \unix\ Network File System (NFS). (2) Our focus is not
exactly on Google's Mapreduce paradigm~\cite{mapreduce} that works on
association lists, but we also provide an API similar to Google's that
works on association lists, just to satisfy popular interest.

%TODO
data locality: scheduling supplied by the user

speeding up the end of computation with duplicate tasks on idle workers
(straightforward since already done for fault tolerance)

@@

\acks

Acknowledgments, if needed.

>>>>

\newcommand{\Ocaml}{OCaml}
\newcommand{\JoCaml}{Jo{\&\!}Caml}
\newcommand{\unix}{\textsc{Unix}}

@@

\begin{abstract}

\end{abstract}

\category{D1.1}{Programming techniques}{Applicative (Functional) Programming}

\terms
Distributed Computing, Functional Programming

%\keywords

@@

\section{Conclusion and Future Work}\label{sec:future}

We presented a distributed programming environment for functional
programming. The main features are the genericity of the interface,
which makes use of polymorphic higher-order functions, and the ability
to easily switch between sequential, multi-cores and network
implementations. A distinguishing feature of our library is also a
robust fault tolerance mechanism implemented within, which relieves
the user from cumbersome implementation details.

Closest to the approach in this paper is Yohann Padioleau's mapreduce
implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
top of OcamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing.  On one hand, we have less flexibility
w.r.t. deployment of the user program; on the other hand, we provide a
more generic API together with fault tolerance.
There are other distributed computing libraries on top of which one
could implement the library discussed in this paper. \JoCaml\
is one of them~\cite{jocaml}. However, \JoCaml\ does not allow
arbitrary closures to be passed between machines, which would limit
the genericity of the API. Additionally, \JoCaml\ does not provide
fault tolerance, which is essential in a distributed setting.
% TODO check with Louis

There are other implementations of distributed computing in the
context of functional programming. One is the Disco
project~\cite{disco}, which implements exactly Google's mapreduce in
Erlang~\cite{erlang}. Our library, on the contrary, is not an \Ocaml\
implementation of Google's mapreduce; rather it is more generic and
not restricted to just key/value pairs.
There are other ways to exploit multi-core architectures. One of which
is data parallelism, which is also relevant in the functional
programming setting~\cite{parallel-haskell}. Our work
does not target data parallelism at all.

There are still some interesting features that could be added to our
library. 
\begin{itemize}
\item 
  One is the ability to efficiently assign tasks to workers depending
  on resource parameters, such as data locality, CPU power, memory,
  etc. This could be achieved by providing the user ways to control
  task scheduling.  
\item 
  Another interesting feature could be the ability
  to add or remove workers dynamically. Currently, our library assumes
  the list of workers to be used is given \emph{a priori}.
\item
  Our library provides limited support for displaying real-time
  information about the computations. Processing and storing
  information about tasks locally in the master is straightforward; 
  monitoring it in real-time is less obvious.
\item 
  One very nice feature of Google's mapreduce is the possibility to
  use idle workers on duplicated tasks to speedup the end of computation.
  Since we already have the fault tolerance implemented, this
  optimization should be straightforward to add to our library.
\end{itemize}
We intend to enrich our library with all above features.

@@

% \acks

% Acknowledgments, if needed.
